---
title: "Music Harvesting"
author: "Álvaro Sanz, Alejandro Aísa"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction 

Music nowadays is probably one of the most profitable industries in the world; almost everyone listens to music, whether it is while driving, when doing exercise, or just lying in bed. This popularity of music has induced to the development huge number of genres, each one with their own style and characteristics. Similarly, the development of the telecommunication technologies in the last couples of decades has made possible that the channels for listening to music change over time. We are not any more confined to CDs and radio. Streaming platforms like Spotify or Video platforms like YouTube are considered nowadays the most important ones for the industry. 

Therefore, the main objective of this work would be to analyse the differences in music. Using the Spotify Web API as a reference, we will try to analise the differences in popularity, genres or characteristics. In this line of reasoning, we will also evaluate the change of music over time. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(httr2)
library(httr)
library(ggplot2)
library(jsonlite)
library(xml2)
library(stringr)
library(stringdist)
library(stargazer)
library(plotly)
```

## Setting the Spotify API Credential 

### Creating the account and the app for developers. 

As explained in the README file, the first necessary conditions for obtaining the credentials for using the Spotify Web API is having a valid account for the [application](https://www.spotify.com). 

Once we have our account, we are ready to register an app within the API; by clicking in the create and app button, and provide a name for the app and a description. The purpose of creating this app is obtaining two central things: the _client ID_ and the _client secret_. These two 

#### Client ID and Client Secret 

Once we have created our account and our Spotify App for developers, we will have to look for our Client ID and Client Secret. They are shown in the dashboard of the App: 

```{r}
client_ID <- scan("C:/Users/alvar/Desktop/client_id.txt", what = character()) # Enter the correct path where the client_id is located. Note: in R, the "\" symbol is not supported. Instead, use "/".
```


```{r}
client_secret <- scan("C:/Users/alvar/Desktop/client_secret.txt", what = character()) # Enter the correct path where the client_secret is located. Note: in R, the "\" symbol is not supported. Instead, use "/".
```

### Creating the personal OAth 2.0 token for requests 

Once we have our client id and our Client Secret, all we have to do is to make a request to the API to obtain the token that would allow us to perform future requests. As a note, this request would be done with httr library, instead of hhtr2. Therefore: 

```{r}
token_req <- POST(
  "https://accounts.spotify.com/api/token",
  accept_json(),
  authenticate(client_ID, client_secret), 
  body = list(grant_type = 'client_credentials'),
  encode = 'form',
  verbose())
```

Once the request is done, we will need to extract the token from the body of the response, and store it in the environment. 

```{r}
mytoken <- content(token_req)$access_token
HeaderValue <- paste0("Bearer ", mytoken)
```

## The Data Harvesting process

### Step 0: setting the base request.

The first task that we should is to define the base URL for future requests. As the endpoints within the API do not possess required fields*, we would be able to construct the queries based on the following URL:

```{r}
spo_gen <- "https://api.spotify.com/v1"
```

Similarly, for practical purposes, we will define now the structure of the request. First, we add the _request_ function of HTTR2 package with the base URL. Then, we provide the token as a personal identification. Finally, as we would not like to perform many requests in a small fraction of time, we include a timeout and a throttle between each one. 

```{r}
req <- request(spo_gen) %>% 
  req_auth_bearer_token(mytoken) %>% 
  req_retry(max_tries = 3) %>% 
  req_options(timeout_ms = 10000) %>% 
  req_throttle(rate = 30 / 60)
```

### 1. Comparing songs in the Global Top 50.

The first request that me would perform to the Spotify Web API involves the playlist with the 50 most listened songs the last day. As an introductory note, each playlist, author or track have their own id, which in URL terms is composed by a series of numbers and letter. This identification string would be appended to the base URL of the beginning and the particular endpoint. Thus, as a first step in the queries, we would need to define the id of the desired playlist/author/track: 

```{r}
top50global <- "37i9dQZEVXbNG2KDcFcKOF?si"
```

Thus, the request is done via the req function defined before. After the request is defines, using the req_url_path_append function we paste the endpoint necessary and the id. Finally, all we have to do is to perform the request and ask for JSON string as the body of the response. Notice here the simplifyVector = T in order to get the key:value pairs stored as dataframes. This structure of the request would be standardized over the following analysis. 

```{r}
top50 <- req %>% 
  req_url_path_append(paste("playlists", top50global, sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = T)

top50
```

Even with the help of simplifyVector = T we have obtain a named list with the values of tracks, ad two dataframes. The second one include other dataframes within each row. Therefore, the main strategy to follow is to unnest the information embedded in this structure. Luckily, we know exactly which value are we looking for, so using the unnest strategy we may obtain the information we want: the names of the song, the duration, whether they are explicit... 

Using the unnest function, we are able to construct a dataframe out of the lists included in the track column. The resulting dataframe is also composed by list columns. Therefore, we would require to perform a second unnest function to finally obtain the desired dataframe with the information about each song of the top50 playlist: 

```{r}
top50 <- top50$tracks$items

top50 <- top50 %>% 
  select(track) %>% 
  unnest(track) %>% 
  select(c(album, artists, duration_ms, explicit, name, id, popularity)) 
```


```{r}
top50filtered <- top50 %>% unnest(artists, names_repair = "universal") %>% #Featurings make 2 different cases, for each artist. 
  select(!c(external_urls, href, type, uri, album))

top50filtered
```

For more precise analysis - each song's features: 
```{r}
songs <- top50filtered$id...11
```

```{r}
req %>% 
  req_url_path_append(paste("audio-features", songs[1], sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = TRUE) %>% 
  as_tibble()
```

```{r}
top50songs <- data.frame()

for(i in 1:length(unique(songs))) { #We set the unique to avoid the features problem. 
  
  song <- req %>% 
  req_url_path_append(paste("audio-features", unique(songs)[i], sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = TRUE) %>% 
  as.tibble()
  
  top50songs <- rbind(song, top50songs)
  
} 

top50songs
```

Now, bind a column for each song's name. 

```{r}
top50ids <- top50filtered %>% 
  select(c(name...10, id...11)) %>% 
  distinct(name...10, .keep_all = TRUE) %>% 
  rename("id" = "id...11")
```

```{r}
top50full <- top50songs %>% 
  full_join(top50ids, by = "id") %>% 
  select(c(`name...10`, danceability, energy, loudness, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_ms))

top50full
```

The tracks appear in inverse order respective to their position in the Top 50, so let's fix that and add that info: 

```{r}
top50full <- top50full %>% 
  arrange(desc(row_number())) %>% 
  mutate(Position = row_number(),
         duration_ms = duration_ms / 1000) %>% 
  rename("duration" = "duration_ms")

top50full
```

How do this 'internal' characteristics of the songs affect their position? Is there any pattern?

```{r}
pois <- glm(Position ~ danceability + energy + loudness + speechiness + acousticness + instrumentalness + liveness + valence + tempo + duration, family = poisson(), data = top50full)

summary(pois)

stargazer(pois,
          type = "text", 
          column.labels = "Coefficient")
```
(Explain)

```{r}
poisson_plot <- function(model) {
  # Extract the model coefficients and standard errors
  coef_df <- data.frame(coef = coef(model), se = sqrt(diag(vcov(model))))
  
  # Plot the coefficient estimates with confidence intervals
  ggplot(coef_df, aes(x = rownames(coef_df), y = coef)) +
    geom_point() +
    geom_errorbar(aes(ymin = coef - 1.96 * se, ymax = coef + 1.96 * se), width = 0.2) +
    coord_flip() +
    xlab("") +
    ylab("Coefficient Estimate") +
    ggtitle("Poisson Regression Coefficient Estimates")
}

poisson_plot(pois)
```


Now, let's visualize if there are notable differences between the top songs and the bottom ones: 

Danceability: 
```{r}
dance <- top50full %>% 
  ggplot(aes(x=Position, y=danceability, color = danceability, text = paste0("Song: ", name...10))) +
  geom_point() +
  ylab("Danceability") +
  xlab("Position in the TOP 50") +
  scale_color_gradient(low = "darkblue", high = "orange") +
  ylim(c(0.25, 1)) +
  theme_light() +
  labs(title = "If I can't dance...",
       color = "Danceability") +
  theme(text = element_text(family = "Helvetica"),
        legend.text = element_text(family = "Helvetica"))

ggplotly(dance, tooltip = c("text", "color")) %>% 
  layout(annotations = list(
  text = "Relationship between the Global position in the Top 50 songs and the degree of danceability of each one",
  x = 0,
  y = 1.05,
  xref = "paper",
  yref = "paper",
  showarrow = FALSE
))
```

Acousticness: 
```{r}

acoustic <- top50full %>% 
  ggplot(aes(x=Position, y=acousticness, color = acousticness, text = paste0("Song: ", name...10))) +
  geom_point() +
  ylab("Speechiness") +
  xlab("Position in the TOP 50") +
  scale_color_gradient(low = "darkblue", high = "orange") +
  ylim(c(0, 1)) +
  theme_light() +
  labs(title = "What about the acoustics?",
       color = "Acousticness") +
  theme(text = element_text(family = "Helvetica"),
        legend.text = element_text(family = "Helvetica"))

ggplotly(acoustic, tooltip = c("text", "color")) %>% 
  layout(annotations = list(
  text = "Relationship between the Global position in the Top 50 songs and the degree of acousticness of each one",
  x = 0,
  y = 1.05,
  xref = "paper",
  yref = "paper",
  showarrow = FALSE
))
```
Speechiness: 

```{r}


speech <- top50full %>% 
  ggplot(aes(x=Position, y=speechiness, color = speechiness, text = paste0("Song: ", name...10))) +
  geom_point() +
  ylab("Speechiness") +
  xlab("Position in the TOP 50") +
  scale_color_gradient(low = "darkblue", high = "orange") +
  ylim(c(0, 0.5)) +
  theme_light() +
  labs(title = "You left me wordless!",
       color = "Speechiness") +
  theme(text = element_text(family = "Helvetica"),
        legend.text = element_text(family = "Helvetica"))

ggplotly(speech, tooltip = c("text", "color")) %>% 
  layout(annotations = list(
  text = "Relationship between the Global position in the Top 50 songs and the density of words in each one",
  x = 0,
  y = 1.05,
  xref = "paper",
  yref = "paper",
  showarrow = FALSE
))
```


### 2. Comparing top songs of different countries

```{r}
top50colombia <- "37i9dQZEVXbL1Fl8vdBUba"
top50india <- "37i9dQZEVXbMWDif5SCBJq"
```
    
#### 2.1. Colombia    

```{r}
top50col <- req %>% 
  req_url_path_append(paste("playlists", top50colombia, sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = T)

top50col <- top50col$tracks$items %>% 
  select(track) %>% 
  unnest() %>% 
  select(c(artists, explicit, duration_ms, name, id, popularity))

colsongs <- top50col$id
```

```{r}
top50colsongs <- data.frame()

for(i in 1:length(unique(colsongs))) { #We set the unique to avoid the features problem. 
  
  song <- req %>% 
  req_url_path_append(paste("audio-features", colsongs[i], sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = TRUE) %>% 
  as.tibble()
  
  top50colsongs <- rbind(song, top50colsongs)
  
} 

top50colsongs
```

```{r}
top50col <- top50col %>% 
  select(c(name, id))
```

```{r}
top50colfull <- top50col %>% 
  full_join(top50colsongs, by = "id") %>% 
  mutate(Position = row_number(),
         playlist = "Colombia TOP 50") %>% 
  select(c(name, Position, danceability, energy, key, loudness, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_ms, playlist))

top50colfull
```

#### 2.2. India

```{r}
top50ind <- req %>% 
  req_url_path_append(paste("playlists", top50india, sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = T)

top50ind <- top50ind$tracks$items %>% 
  select(track) %>% 
  unnest() %>% 
  select(c(artists, explicit, duration_ms, name, id, popularity))

indsongs <- top50ind$id
```

```{r}
top50indsongs <- data.frame()

for(i in 1:length(unique(indsongs))) { #We set the unique to avoid the features problem. 
  
  song <- req %>% 
  req_url_path_append(paste("audio-features", indsongs[i], sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = TRUE) %>% 
  as.tibble()
  
  top50indsongs <- rbind(song, top50indsongs)
  
} 

top50indsongs
```

```{r}
top50ind <- top50ind %>% 
  select(c(name, id))
```

```{r}
top50indfull <- top50ind %>% 
  full_join(top50indsongs, by = "id") %>% 
  mutate(Position = row_number(), 
         playlist = "India TOP 50") %>% 
  select(c(name, Position, danceability, energy, key, loudness, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_ms, playlist))

top50indfull
```

Are there important differences in the distribution of this variables? 

```{r}
tracks2 <- rbind(top50colfull, top50indfull)
```


```{r}
dance <- ggplot(tracks2, aes(x=danceability, fill=playlist,
                    text = paste(playlist)))+
  geom_density(alpha=0.7, color=NA)+
  scale_fill_manual(values=c("violet", "darkblue"))+
  labs(x="Danceability", y="Density") +
  guides(fill=guide_legend(title="Playlist"))+
  xlim(c(0.25,1)) +
  theme(text = element_text(family = "Helvetica")) +
  theme_minimal()+
  labs(title = "Who dances more?")

ggplotly(dance, tooltip=c("text")) %>% 
  layout(annotations = list(
    text = "How are each countries' songs in terms of danceability?",
    x = 0,
    y = 1.05,
    xref = "paper",
    yref = "paper",
    showarrow = FALSE
  ))
```

```{r}
positive <- ggplot(tracks2, aes(x=valence, fill=playlist,
                    text = paste(playlist)))+
  geom_density(alpha=0.7, color=NA)+
  scale_fill_manual(values=c("violet", "darkblue"))+
  labs(x="Valence", y="Density") +
  guides(fill=guide_legend(title="Playlist"))+
  xlim(c(0.25,1)) +
  theme(text = element_text(family = "Helvetica")) +
  theme_minimal()+
  labs(title = "What a wonderful... song?")

ggplotly(positive, tooltip=c("text")) %>% 
  layout(annotations = list(
    text = "Distribution of songs in terms of positiveness of the songs",
    x = 0,
    y = 1.05,
    xref = "paper",
    yref = "paper",
    showarrow = FALSE
  ))
```

```{r}
speech <- ggplot(tracks2, aes(x=speechiness, fill=playlist,
                    text = paste(playlist)))+
  geom_density(alpha=0.7, color=NA)+
  scale_fill_manual(values=c("violet", "darkblue"))+
  labs(x="Valence", y="Density") +
  guides(fill=guide_legend(title="Playlist"))+
  xlim(c(0,0.5)) +
  theme(text = element_text(family = "Helvetica")) +
  theme_minimal()+
  labs(title = "Who talks more?")

ggplotly(speech, tooltip=c("text")) %>% 
  layout(annotations = list(
    text = "Distribution of songs in terms of density of words",
    x = 0,
    y = 1.05,
    xref = "paper",
    yref = "paper",
    showarrow = FALSE
  ))
```

### 3. Comparing TOP 50s. Spotify Vs. Billboard

```{r}
billboard <- "https://www.billboard.com/charts/billboard-global-200/"
browseURL(billboard)
```

```{r}
billboard_raw <- read_html(billboard) %>% 
  xml_child()
```

Songs: 
```{r}
bill200 <- billboard_raw %>%
  xml_find_all("//li/ul/li/h3[@id = 'title-of-a-story']")
```

```{r}
billfilter <- bill200[1:50] %>% 
  xml_text() %>% 
  str_remove_all("\t|\n")
  
```

Artists:

```{r}
bill200art <- billboard_raw %>%
  xml_find_all("//li/ul/li/span[contains(@class, 'c-label  a-no-trucate a-font-primary')]")
```

```{r}
billfilterart <- bill200art[1:50] %>% 
  xml_text() %>% 
  str_remove_all("\t|\n")
```

Merging both: 
```{r}
top50billboard <- billfilter %>% 
  as.data.frame() %>% 
  cbind(as.data.frame(billfilterart)) %>% 
  rename("Track" = ".",
         "Artist" = "billfilterart") %>% 
  mutate(`Billboard Position` = row_number())

top50billboard
```

And take the positions from the Spotify Global 50: 

```{r}
top50positions <- top50filtered %>% 
  select(c(name...5, name...10)) %>% 
  distinct(name...10, .keep_all = TRUE) %>% 
  mutate(`Spotify Position` = row_number()) %>% 
  rename("Artist" = "name...5", 
         "Track" = "name...10")

top50positions
```

```{r}
top50billboard %>% 
  full_join(top50positions, by = c("Track"))
```

Problem! There are some songs' names that differ slightly from one page to another. How can we fix this?

```{r}
top50billboard$Track_match <- sapply(top50billboard$Track, function(x) top50positions$Track[stringdist(x, top50positions$Track) == min(stringdist(x, top50positions$Track))])
top50positions$Track_match <- sapply(top50positions$Track, function(x) top50billboard$Track[stringdist(x, top50billboard$Track) == min(stringdist(x, top50billboard$Track))])

```

```{r}
top50billboard <- top50billboard %>% 
  unnest()

top50positions <- top50positions %>% 
  unnest()

# full_join the data frames on the matched song names
df_join <- dplyr::full_join(top50positions, top50billboard, by = c("Track_match" = "Track"))
```

```{r}
comparison <- df_join %>% 
  select(c(Artist.x, `Spotify Position`, `Billboard Position`, Track_match)) %>% 
  distinct(Track_match, .keep_all = TRUE) %>% 
  mutate(Difference = `Spotify Position` - `Billboard Position`,
         reference = 0, 
         value = ifelse(Difference > 0, "Positive", ifelse(Difference == 0, "Neutral", "Negative")))

comparison
```

Which are the most penalized songs in comparison?

```{r, fig.height=6}
my_colors <- c("Positive" = "darkgreen", "Neutral" = "gray", "Negative" = "red")

comparison <- comparison %>% 
  na.omit() %>% 
  ggplot(aes(y = reorder(Track_match, -`Spotify Position`), x=reference, color = value)) +
  geom_segment(aes(xend=Difference, yend = Track_match)) + 
  xlim(c(-30, 30)) + 
  xlab("Positions lost/gained in Billboard") +
  ylab("Track") +
  theme_minimal() +
  theme(text = element_text(family = "Helvetica"), 
        panel.grid.minor.x = element_blank()) + 
  scale_color_manual(values = my_colors) +
  geom_point(aes(x= Difference, y=Track_match))
  
ggplotly(comparison, tooltip="Difference")
```


### 4. Comparing music across time 

#### 4.1.  All out X0s

When comparing historic music, we will first resort to the spotify-made playlists for the most famous songs of each decade (60s to 10s. In order to perform this specific request, we will need the specific URI for playlists and the particular IDs from each of them. 

##### Playlists IDs

```{r}
tens <-  "37i9dQZF1DX5Ejj0EkURtP" 
zeros <- "37i9dQZF1DX4o1oenSJRJd"
nineties <-  "37i9dQZF1DXbTxeAdrVG2l" 
eighties <-  "37i9dQZF1DX4UtSsGT1Sbe"
seventies <- "37i9dQZF1DWTJ7xPn4vNaz" 
sixties <- "37i9dQZF1DXaKIA8E7WcJj"
```

For this particular case, we have selected 6 different playlist, one for each decade since the 60s. In order to study their _importance_ for listeners nowadays, we will look for the number of followers that each of them posses. This figure account for the number of people that have stored and downloaded the playlists in their own account. Thus, we may assume this number as a proxy for popularity. 

##### Custom function for extracting followers 

As the data is nested in a JSON within the body of the response we will need to apply again some techniques to read the data into a dataframe. However, contrary to previous points, we will use a function to extract the information in a single step. This custom function will perform the following tasks: 

- In the first place, we will create the specific URI for each playlist 
- Secondly, we will perform the request, provinding the personal token. 
- Next, we will create a dataframe with a column list, for the information provided in the response. Within that dataframe, we will select only the row that contain information about the followers. Then, we will unnest this column to create a new dataframe. 
- The new dataframe for followers contains a list with two values; a NULL and the actual figure. Then, we need to filter out the first and select and rename the colums that pertains to our query. 
- We repeat the previous step for the node that contains the information of the name of the playlist. Lucklily, this time the node only contains such name, so the unnest function serve to create a new one-row dataframe. Lastly, we change the name of the column for practical purposes. 
- Finally, we merge the two dataframes into a sigle one. 

```{r}

followers <- function(x) { 

  resp_output <- 
    req %>% 
    req_url_path_append(paste("playlists", x, sep = "/")) %>% 
    req_auth_bearer_token(mytoken) %>% # Providing the token
    req_perform() %>% # Performing the request
    resp_body_json() # Obtaining the response body as a JSON file 
  
  resp_followers <- 
    resp_output %>% 
    enframe() %>% # Creating the column-list dataframe
    filter(name == "followers") %>% # selecting only the followers node
    unnest(cols = value) %>% # creating the dataframe
    filter(!value == "NULL") %>% # filtering out NULL row 
    select(-name) %>% # Selecting and renaming 
    rename("followers" = "value")
  
  resp_name <- resp_output %>% 
    enframe() %>% # Creating column-list
    filter(name == "name") %>% # Filtering for the name of the playlist
    unnest(cols = value) %>% # Creating the dataframe 
    select(-name) %>% 
    rename("playlist" = "value")
    
  df <- cbind(resp_name, resp_followers) # Merging 
  }
```

Once we have defined our custom function, we will use lapply to create a loop that would go over each individual playlist and extract the information. Then, thanks to the do.call function, we will merge all the figures into a single data frame. 

```{r}
years <- c(sixties, seventies, eighties, nineties, zeros, tens)

followers_list <- lapply(years, followers) 
followers_df <- do.call(rbind, followers_list) %>% transmute(
    playlist = as.character(playlist), 
    followers = as.numeric(followers))
```

##### Plotting the number of followers of each decade playlist. 

As we have now the data stored in a dataframe, we can easily plot them:

```{r}
hp <- ggplot(followers_df)+
  aes(playlist, followers)+
  geom_col(aes(fill = playlist)) + 
  theme_minimal()+
  guides(fill = "none")

hp
```

#### 4.2. Comparing music features on historic music

Similarly, we may create a custom function that would enable us to obtain different musical features of different songs. Spotify indexes different characteristics for each song, such as the danceability, the energy or the tempo. Same as before, we could compare songs/artists from nowadays to older ones, according to this features. Then, we have selected the most liked song (in spotify) from twelve artists, one man and one female from each decade (60s to 10s). For function's purposes, we will store them already in a vector. Also, we will store the particular URI por the tracks' endopoint:  Their IDs in spotify are the following: 

```{r}
Rihanna <- "49FYlytm3dAAraYgpoJZux"
Drake <- "1zi7xx7UVEFkmKfv06H8x0"
Eminem <- "1v7L65Lzy0j0vdpRjJewt1"
LadyGaga <- "1QV6tiMFM6fSOKOGLMHYYg"
MichaelJackson <- "3S2R0EVwBSAVMd5UMgKTL0"
Madonna <- "22sLuJYcvZOSoLLRYev1s5"
Queen <- "3z8h0TU7ReDPLIbEnYhWZb"
Abba <- "0GjEhVFGZW8afUYGChu3Rr"
RollingStones <- "63T7DJ1AFDD6Bn8VzG6JE8"
WhitneyHouston <- "2tUBqZG2AbRi7Q0BIrVrEj"
TheBeatles <- "6dGnYIeXmHdcikdzNNDMm2"
Cher <- "2goLsvvODILDzeeiT4dAoR"

artists <- c(Rihanna, Drake, Eminem, LadyGaga, 
            MichaelJackson, Madonna, Queen, Abba, 
            RollingStones, WhitneyHouston, TheBeatles, Cher,
            TheBeatles, Cher)
```


##### Custom function 

The JSON returned in the request of the features is less nested. Therefore, the function is simpler. Similarly, we may use _simplifyVector = T_ and _as.tibble_ to have the key:value already stored as dataframes. Thus, the function will bear the request URI, the OAth token and the storing specificities.

```{r}
features <- function(x) {

  req <- 
    req %>% 
    req_url_path_append(paste("audio-features", x, sep = "/")) %>%  
    req_auth_bearer_token(mytoken) %>% 
    req_perform() %>% 
    resp_body_json(simplifyVector = T) %>% 
    as_tibble() 
  
}
```

Once we have the feature function, we will use _lapply_ and _do.call_ to perform the request for all the tracks and bind them together in a dataframe. For visualizing purposes, we will add a new colum specifiying the author of the particular songs: 

```{r}
features_list <- lapply(artists, features)

features_df <- do.call(rbind, features_list) 

features_df <- features_df %>% 
  mutate(
    author = case_when(
      id == "49FYlytm3dAAraYgpoJZux" ~ "Rihanna", 
      id == "1zi7xx7UVEFkmKfv06H8x0" ~ "Drake", 
      id == "1v7L65Lzy0j0vdpRjJewt1" ~ "Eminem", 
      id == "1QV6tiMFM6fSOKOGLMHYYg" ~ "Lady Gaga", 
      id == "3S2R0EVwBSAVMd5UMgKTL0" ~ "Michael Jackson", 
      id == "22sLuJYcvZOSoLLRYev1s5" ~ "Madonna", 
      id == "3z8h0TU7ReDPLIbEnYhWZb" ~ "Queen", 
      id == "0GjEhVFGZW8afUYGChu3Rr" ~ "Abba", 
      id == "63T7DJ1AFDD6Bn8VzG6JE8" ~ "Rolling Stone", 
      id == "2tUBqZG2AbRi7Q0BIrVrEj" ~ "Whitney Houston", 
      id == "6dGnYIeXmHdcikdzNNDMm2" ~ "Beatles", 
      id == "2goLsvvODILDzeeiT4dAoR" ~ "Cher")) %>% 
  dplyr::select(author, danceability, energy, key, loudness, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_ms) %>% 
    mutate(
      decade = case_when(
        author == "Rihanna" |  author == "Drake" ~ "10s", 
        author == "Lady Gaga" |  author == "Eminem" ~ "00s", 
        author == "Madonna" |  author == "Michael Jackson" ~ "90s", 
        author == "Abba" |  author == "Queen" ~ "80s", 
        author == "Whitney Houston" |  author == "Rolling Stone" ~ "70s", 
        author == "Cher" |  author == "Beatles" ~ "60s")) 
        
    
```

##### Plotting the features

```{r}
features_plot <- ggplot(features_df) +
  aes(x = danceability, y = energy, colour = decade) +
  geom_point() +
  scale_size_area(max_size = 20)+
  theme_minimal()+
  geom_text(aes(label = author)) 

features_plot

```

#### 4.3. Combining features from historic songs

As a final step, we may combine the utilities of the custom functions, the historical All Out playlist, and the analysis of the musical features to analyse the differences in styles for the music for the different decades. 

##### Custom function

The first task consists in obtaining all the ids of the All Out playlist, to look for their features. For that, we will use the third and final custom function follows the same logic as the previous one. First, we construct the request, by adding the endpoint. Then, once we have the JSON for the playlist information, we dig in into the nested key:value pairs until we get the vector with the ids. Also, for identification purposes, we rbind the id to the name of the list

```{r}
features <- function(x) {

all_out <- 
  req %>% 
  req_url_path_append(paste("playlists", x, sep = "/")) %>% 
  req_perform() %>% # Performing the request
  resp_body_json(simplifyVector = T)

all_out_ids = all_out$tracks$items %>% 
  dplyr::select(track) %>% 
  unnest(track) %>% 
  dplyr::select(id)

name <- all_out$name

df <- cbind(all_out_ids, name)

}
```

Once we constructed the function, we can use lapply and do.call to run the fucntion for all the All Outs playlists and store them in a dataframe.

```{r}
features_list <- lapply(years, features)
features_df <- do.call(rbind, features_list)
```

We will select only a sample of song, so we don't overload the request at the same time we are not having the computing rendering for a lot of time (taking into account the timeout and the throttle)

```{r}
tracks_AO <- sample(features_df$id, 100)
```

Then, we apply the loop created before for storing the information into a dataset. 

```{r}
AOsongs <- data.frame()

for(i in 1:length(unique(tracks_AO))) { #We set the unique to avoid the features problem. 
  
  song <- req %>% 
    req_url_path_append(paste("audio-features", unique(tracks_AO)[i], sep = "/")) %>% 
    req_perform() %>% 
    resp_body_json(simplifyVector = TRUE) %>% 
    as_tibble()
  
  AOsongs <- rbind(song, AOsongs)
  
}

AOsongs
```
We then remove the columns that do not add relevant information and rename the column of the name of the playlist.  

```{r}
AO_Full <- AOsongs %>% 
  dplyr::select(-uri, -track_href, -analysis_url, -type) %>% 
  left_join(features_df, by = "id") %>% 
  dplyr::rename("playlist" = "name")
```
The first visualization we may do with the set of data is to compare the energy transmitted by a song and it danceability: 

```{r}
featuresAO_plot <- ggplot(AO_Full) +
  aes(x = energy, y = danceability, colour = playlist) +
  geom_point() +
  scale_size_area(max_size = 20)+
  theme_minimal()

plotly::ggplotly(featuresAO_plot, tooltip=c("playlist"))
```

When plotting the two variables together we find that there seem to be no linear relationship between danceability and energy. However, we do se that there is a tendency to be in the right part of the plot. 

Finally, using the same code as before for the danceability in the top50, we can plot the energy transmitted by decade. Also, apart from changing the variable, we will also add a facet wrap to visualize

```{r}
viz5 <- ggplot(AO_Full, aes(x=energy, fill=playlist,
                    text = paste(playlist)))+
  geom_density(alpha=0.7, color=NA)+
  scale_fill_manual(values=c("blue", "orange", "pink", "green", "grey", "red"))+
  labs(x="Energy", y="Density") +
  guides(fill=guide_legend(title="Playlist"))+
  xlim(c(0.25,1)) +
  theme_minimal()+
  ggtitle("Energy") +
  facet_wrap(~playlist, ncol = 3)

plotly::ggplotly(viz5, tooltip=c("text"))
```

The plot shows the huge differences between decades. The songs of the 60s are indeed distributed in the first part of the scale, meaning that they could in fact be less energetic than the rest. Interestengly, for the song of the 00s decade there is a minimum amount of energy of the song:













