---
title: "Working Rmd"
author: '100385774'
date: "2023-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction and objectives
Alvaro: 1-4
Alex: 5-x
1. Compare the 50 TOP GLOBAL SONGS (weekly playlist) - features: https://open.spotify.com/playlist/37i9dQZEVXbNG2KDcFcKOF?si=3a61996b79fd4698 
2. Compare 2 different countries WEEKLY TOP 50 - features: 
    - Colombia: https://open.spotify.com/playlist/37i9dQZEVXbL1Fl8vdBUba?si=6c0379f1a546446c
    - India: https://open.spotify.com/playlist/37i9dQZEVXbMWDif5SCBJq?si=becf0ad688bc4894
3. Compare TOP 50 from Billboard (Web scraping) vs. TOP 50 from Spotify - track/artists differences
4. Correlate TOP 50 GLOBAL popularity with YouTube views (web scraping) 
5. Variability between GLOBAL DAILY TOP 50 in a week - tracks/artists: https://open.spotify.com/playlist/37i9dQZEVXbMDoHDwVN2tF?si=05a703094b4841be
6. TOP SONGS evolution over decades -  
7. Artist's evolution over time - 

# Scraping

```{r}
library(tidyverse)
library(httr2)
library(httr)
library(ggplot2)
library(jsonlite)
library(xml2)
library(stringr)

set_config(
  user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64; rv:109.0) Gecko/20100101 Firefox/109.0; Alvaro Sanz / 100385774@alumnos.uc3m.es")
)
```


## API authentication

### Set the credentials

```{r}
CLIENT_ID <- "d1552a9b168446f49583043fed2ae16e"
CLIENT_SECRET <- "baeda3593af24234a8aea32e5c22fa48"
```

### Setting the endpoints

```{r}

spo_gen <- "https://api.spotify.com/v1"
spo_token <- "https://accounts.spotify.com/api/token"

```

### Token try httr

```{r}
token1 <-  POST("https://accounts.spotify.com/api/token",
  accept_json(),
  authenticate(CLIENT_ID, CLIENT_SECRET), 
  body = list(grant_type = 'client_credentials'),
  encode = 'form',
  verbose()
)

token1

fulltoken <-  content(token1)$access_token
```

### Base req settings:
```{r}
req <- request(spo_gen) %>% 
  req_auth_bearer_token(fulltoken) %>% 
  req_retry(max_tries = 3) %>% 
  req_options(timeout_ms = 10000) %>% 
  req_throttle(rate = 30 / 60)
```

# 1. Compare the 50 TOP GLOBAL SONGS (weekly playlist) - features: https://open.spotify.com/playlist/37i9dQZEVXbNG2KDcFcKOF?si=3a61996b79fd4698 

```{r}
top50global <- "37i9dQZEVXbNG2KDcFcKOF?si"
```


```{r}
top50 <- req %>% 
  req_url_path_append(paste("playlists", top50global, sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = T)

top50

top50 <- top50$tracks$items

top50 <- top50 %>% 
  select(track) %>% 
  unnest(track) %>% 
  select(c(album, artists, duration_ms, explicit, name, id, popularity)) 
```

```{r}
top50filtered <- top50 %>% unnest(artists, names_repair = "universal") %>% #Featurings make 2 different cases, for each artist. 
  select(!c(external_urls, href, type, uri, album))

top50filtered
```
For more precise analysis - each song's features: 
```{r}
songs <- top50filtered$id...11
```

```{r}

req %>% 
  req_url_path_append(paste("audio-features", songs[1], sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = TRUE) %>% 
  as.tibble()

```

```{r}

top50songs <- data.frame()

for(i in 1:length(unique(songs))) { #We set the unique to avoid the features problem. 
  
  song <- req %>% 
  req_url_path_append(paste("audio-features", unique(songs)[i], sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = TRUE) %>% 
  as.tibble()
  
  top50songs <- rbind(song, top50songs)
  
} 

top50songs
```

Now, bind a column for each song's name. 

```{r}
top50ids <- top50filtered %>% 
  select(c(name...10, id...11)) %>% 
  distinct(name...10, .keep_all = TRUE) %>% 
  rename("id" = "id...11")
```

```{r}
top50full <- top50songs %>% 
  full_join(top50ids, by = "id")

top50full
```


2. Compare 2 different countries WEEKLY TOP 50 - features: 
    - Colombia: https://open.spotify.com/playlist/37i9dQZEVXbL1Fl8vdBUba?si=6c0379f1a546446c
    - India: https://open.spotify.com/playlist/37i9dQZEVXbMWDif5SCBJq?si=57eaf0528d76483e
    
```{r}
top50colombia <- "37i9dQZEVXbL1Fl8vdBUba"
top50india <- "37i9dQZEVXbMWDif5SCBJq"
```
    
## Colombia    

```{r}
top50col <- req %>% 
  req_url_path_append(paste("playlists", top50colombia, sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = T)

top50col <- top50col$tracks$items %>% 
  select(track) %>% 
  unnest() %>% 
  select(c(artists, explicit, duration_ms, name, id, popularity))

colsongs <- top50col$id
```

```{r}
top50colsongs <- data.frame()

for(i in 1:length(unique(colsongs))) { #We set the unique to avoid the features problem. 
  
  song <- req %>% 
  req_url_path_append(paste("audio-features", colsongs[i], sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = TRUE) %>% 
  as.tibble()
  
  top50colsongs <- rbind(song, top50colsongs)
  
} 

top50colsongs
```

```{r}
top50col <- top50col %>% 
  select(c(name, id))
```

```{r}
top50colfull <- top50col %>% 
  full_join(top50colsongs, by = "id")

top50colfull
```

## India

```{r}
top50ind <- req %>% 
  req_url_path_append(paste("playlists", top50india, sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = T)

top50ind <- top50ind$tracks$items %>% 
  select(track) %>% 
  unnest() %>% 
  select(c(artists, explicit, duration_ms, name, id, popularity))

indsongs <- top50ind$id
```

```{r}
top50indsongs <- data.frame()

for(i in 1:length(unique(indsongs))) { #We set the unique to avoid the features problem. 
  
  song <- req %>% 
  req_url_path_append(paste("audio-features", indsongs[i], sep = "/")) %>% 
  req_perform() %>% 
  resp_body_json(simplifyVector = TRUE) %>% 
  as.tibble()
  
  top50indsongs <- rbind(song, top50indsongs)
  
} 

top50indsongs
```

```{r}
top50ind <- top50ind %>% 
  select(c(name, id))
```

```{r}
top50indfull <- top50ind %>% 
  full_join(top50indsongs, by = "id")

top50indfull
```

3. Compare TOP 50 from Billboard (Web scraping) vs. TOP 50 from Spotify - track/artists differences


```{r}

billboard <- "https://www.billboard.com/charts/billboard-global-200/"
browseURL(billboard)
```

```{r}
billboard_raw <- read_html(billboard) %>% 
  xml_child()
```

Songs: 
```{r}
bill200 <- billboard_raw %>%
  xml_find_all("//li/ul/li/h3[@id = 'title-of-a-story']")
```

```{r}
billfilter <- bill200[1:50] %>% 
  xml_text() %>% 
  str_remove_all("\t|\n")
  
```

Artists:

```{r}
bill200art <- billboard_raw %>%
  xml_find_all("//li/ul/li/span[contains(@class, 'c-label  a-no-trucate a-font-primary')]")
```

```{r}
billfilterart <- bill200art[1:50] %>% 
  xml_text() %>% 
  str_remove_all("\t|\n")
```

Merging both: 
```{r}
billfilter %>% 
  as.data.frame() %>% 
  cbind(as.data.frame(billfilterart)) %>% 
  rename("Artist" = ".",
         "Track" = "billfilterart") %>% 
  mutate(position = row_number())
```








